---
title: "RIOS Research - Inclusive Teaching Text Analysis"
author: "Sokona Mangane"
output: html_document
date: "2022-12-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup/Data Cleaning

Here I'm installing and loading the necessary packages and doing some data cleaning. 

```{r Loading packages and data}
#install.packages("skimr")
#install.packages("tidyverse")
#install.packages("tidytext")
#install.packages("stopwords")
#install.packages("wordcloud")
#install.packages("reshape2")
#install.packages("ggraph")
library(ggraph)
library(igraph)
library(skimr)
library(tidyverse)
library(tidytext)
library(ggplot2)
library(readr)
library(stopwords)
library(wordcloud)
library(reshape2)
rios_data <- read_csv("RIOS Research - Course Source - Sheet1 2.csv")
dei_keywords <- read_csv("SJEDI_words 2022-12-20 18_03_42.csv")
rios_data$`Inclusive Teaching  included?`[12] = "No"
rios_data <- rios_data %>%
  arrange(desc(Year))
```

I've created a variable with diversity related words from the keywords column and then added to the DEI Keywords that I imported `dei_keywords`. I also added another column, which includes the article number for each row.

``` {r More Data Cleaning}
#diversity related words from the 'keyword themes' column!
diversity_related <- c("diversity", "bias", "confirmation bias", "cognitive bias", "social justice", "broader impacts", "racism", "identity", "equity", "inclusivity", "environmental justice", "inclusion", "belonging")

#adding the vector above to the CSV dei_keywords
for (x in 1:13){
  dei_keywords[nrow(dei_keywords) + 1,] = diversity_related[x]
}

rios_data$article_num <- c(1:nrow(rios_data))
```

Here, each word from the `Inlcusive Teaching Description` and `keyword themes` is "un-nested" into it's own row. Then any unnecessary punctuation, numbers and words are removed, for analysis.

``` {r More Data Cleaning 2}

rios_data_tokenizedit <- rios_data %>%
  unnest_tokens(output = inclusive_teach_tokens, input = `Inclusive Teaching Description`)


#removing all rows with any punctuation, digits, or "stopwords" (~20k rows total)
strings <- c("[:punct:]", "[:digit:]","\\(","\\)")
stopwords_vec <- stopwords(language = "en")
stopwords_vec <- stopwords_vec[-c(165:167)]

#removed ~777 rows
rios_data_tokenizedit <- rios_data_tokenizedit %>%
  filter(!str_detect(inclusive_teach_tokens, paste(strings, collapse = "|")))

#removed ~19,663 rows
rios_data_tokenizedit <- rios_data_tokenizedit %>%
  filter(!inclusive_teach_tokens %in% stopwords_vec) 

#doing same thing as above but for keyword themes
rios_data_tokenizedkt <- rios_data %>%
  unnest_tokens(output = keyword_themes_tokens, input = `keyword themes`)


#removing all rows with any punctuation, digits, or "stopwords" (78 rows total)
strings <- c("[:punct:]", "[:digit:]","\\(","\\)")
stopwords_vec <- stopwords(language = "en")

#removed ~777 rows
rios_data_tokenizedkt <- rios_data_tokenizedkt %>%
  filter(!str_detect(keyword_themes_tokens, paste(strings, collapse = "|")))

#removed ~19,663 rows
rios_data_tokenizedkt <- rios_data_tokenizedkt %>%
  filter(!keyword_themes_tokens %in% stopwords_vec) 


```

## EDA

`dei_related` columns were created for each data frame, which says TRUE if that word (regarding the Inclusive Teaching Descriptions and the keyword themes) matches any word from the `dei_keywords` dataframe.

```{r EDA}

#creating a DEI related column
rios_data_tokenizedit$dei_relatedit = NA

rios_data_tokenizedkt$dei_relatedkt = NA

rios_data_tokenizedit$dei_relatedit <- sapply(rios_data_tokenizedit$inclusive_teach_tokens, function(x) any(sapply(dei_keywords, str_detect, string = x)))

rios_data_tokenizedkt$dei_relatedkt <- sapply(rios_data_tokenizedkt$keyword_themes_tokens, function(x) any(sapply(dei_keywords, str_detect, string = x)))

#save(dei_keywords, file = "dei_keywords.csv")
#saveRDS(rios_data_tokenized, file = "rios_data_tokenized.rds")

#removing the unnnecessary columns
rios_data_tokenizedit <- rios_data_tokenizedit[,-c(9:13)]
rios_data_tokenizedkt <- rios_data_tokenizedkt[,-c(9:13)]


#most common DEI words
rios_data_tokenizedit %>%
  filter(dei_relatedit == "TRUE") %>%
  group_by(Year, article_num, inclusive_teach_tokens) %>%
  count(inclusive_teach_tokens, sort = TRUE)

#the most common dei_related words don't necessarily have the same dei related keyword themes
rios_data_tokenizedkt %>%
  filter(dei_relatedkt == "TRUE") %>%
  group_by(Year, article_num, keyword_themes_tokens) %>%
  count(keyword_themes_tokens, sort = TRUE)

#saving for visuals on word counts, etc
it_word_counts <- rios_data_tokenizedit %>%
  filter(dei_relatedit == "TRUE") %>%
  group_by(Year, inclusive_teach_tokens) %>%
  count(inclusive_teach_tokens, sort = TRUE)

#which articles and years has the highest proportion of DEI related words?
#find the # of dei related words in each article/year and divide by the total words in each article/year

#totaldei for each year
totaldeiyear <- it_word_counts %>%
  group_by(Year) %>%
  summarise(totaldeiy = sum(n))

#totaldei for each article, only 251 obs, so some articles don't have any dei related words (30 articles with no Inclusive Teaching Section...)
totaldeiarticle <- rios_data_tokenizedit %>%
  filter(dei_relatedit == "TRUE") %>%
  group_by(article_num) %>%
  count(inclusive_teach_tokens) %>%
  summarise(totaldeia = sum(n)) %>%
  arrange(desc(totaldeia))

#total words in each year
totalwordsyear <- rios_data %>%
  group_by(Year) %>%
  summarise(totaly = sum(`Word Count of Inclusive Teaching?`, na.rm = TRUE))

#total words in each article
totalwordsarticle <- rios_data %>%
  select(15,9)
  

#proportions for each year, fluctuating 
totaldeiyear %>%
  full_join(totalwordsyear, by = c("Year")) %>%
  mutate(ratio = totaldeiy/totaly)%>%
  arrange(desc(ratio))

#proportions for each article, also fluctuating (higher article # means it happened earlier ex: article # 270 = year 2014)
totaldeiarticle %>%
  full_join(totalwordsarticle, by = c("article_num")) %>%
  mutate(ratio = totaldeia/`Word Count of Inclusive Teaching?`) %>%
  arrange(desc(ratio))



```

## Visuals :)


```{r}
#word count over time plot 
#rios_data %>%
 # ggplot(aes(rios_data$Year, rios_data$`Word Count of Inclusive Teaching?`)) + 
  #  geom_boxplot(na.rm = TRUE) +
   # labs(x = "Year", y = "Word count of Inclusive Teaching Section",
  #title = "Word Count of Inlcusive Teaching Sections Over Time")


  boxplot(`Word Count of Inclusive Teaching?`~Year,
          data=rios_data,
          main="Word Count of Inlcusive Teaching Sections Over Time",
          xlab="Year",
          ylab="Word count of Inclusive Teaching Section",
          horizontal = TRUE)



```

The x axis is very crowded!

```{r plot time!, warning=FALSE}

#it_word_counts is defined in chunk 4!

it_word_counts %>%
  top_n(20) %>%
  mutate(inclusive_teach_tokens = reorder(inclusive_teach_tokens, n)) %>%
  ggplot(aes(inclusive_teach_tokens, n, fill = Year)) + 
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~Year, ncol = 2, scales = "free_x")
```

Better (w/o facet wrap), but only includes ~22% of data! (b/c we filtered for n > 5)

``` {r}
it_word_counts %>%
  filter(n > 5) %>%
  mutate(inclusive_teach_tokens = reorder(inclusive_teach_tokens, n)) %>%
  ggplot(aes(inclusive_teach_tokens, n, fill = Year)) +
  geom_col() +
  coord_flip() +
  labs(y = "(DEI Related) Word Count in Inclusive Teaching Text") +
  facet_wrap(~Year)
```

Cool! but why are some words showing up twice? (ex.. identities, engage, abilities, diverse)

``` {r}
it_word_counts %>%
  with(wordcloud(inclusive_teach_tokens, n))
```

??? Too many years! and not color blind friendly

``` {r}
it_word_counts %>%
  acast(inclusive_teach_tokens ~ Year, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("#fbb4ae", "#b3cde3", "#ccebc5", "#decbe4", "#fed9a6", "#ffffcc", "#e5d8bd", "#fddaec", "#f2f2f2"))

```

## 2-gram Analysis

```{r 2-gra, word analysis}


rios_data_token2it <- rios_data %>%
  unnest_tokens(it_tokens_2w, `Inclusive Teaching Description`, token = "ngrams", n = 2)  %>%
  separate(it_tokens_2w, c("word1", "word2"), sep = " ")

rios_data_token2it <- rios_data_token2it %>%
  filter(!word1 %in% stopwords_vec) %>%
  filter(!word2 %in% stopwords_vec) %>% 
  unite(it_tokens_2w, word1, word2, sep = " ") 

rios_data_token2it <- rios_data_token2it %>%
  filter(!str_detect(it_tokens_2w, paste(strings, collapse = "|")))


#creating a DEI related column
rios_data_token2it$dei_related = NA


rios_data_token2it$dei_related <- sapply(rios_data_token2it$it_tokens_2w, function(x) any(sapply(dei_keywords, str_detect, string = x)))


#removing the unnnecessary columns/rows
rios_data_token2it <- rios_data_token2it[,-c(9:13)]

unnecessary <- c("individually table", "individual investigators", "individual module", "individual noninteractive", "individual pre", "individual clicker", "inclusion another", "inclusion additionally", "identify species", "identify primer", "identify possible", "identified alternatively", "identification", "ideas individual", "group divide", "general inclusive", "bird identification", "yet inclusive", "variants identified", "teachers identify", "skills perspectives", "sheet individually", "residential birds", "radiation incidents", "regular individually", "pipeline cure", "plant communities", "popular culture", "personal connection", "physical connection", "perspective remind", "perspectives anonymous", "four individuals", "first collaborative", "find identify", "ever identified", "evenly divides", "ethnic economic", "ethnic given", "equity public", "england individual", "engaging final", "diverse face", "diverse natural", "diverse mixed", "direct connection", "disabilities benefit", "data individuals", "data individually", "less confident", "communities due", "collaborative yet", "collaborative easing", "collaboration using", "collaboration throughout", "class individual", "biodiversiy lab", "biodiversity losses", "backgrounds may", "backgrounds find", "backgrounds furthermore", "backgrounds therefore", "area identified", "answer individually", "agricultural sciences", "ability train", "ability moreover", "abilities match", "questions individually", "individuals turn", "individuals since", "communicate collaborate", "backgrounds throughout", "access see", "ada accessibility", "de identified", "efficacy identity", "individualactors may")

rios_data_token2it <- rios_data_token2it %>%
  filter(!str_detect(it_tokens_2w, paste(unnecessary, collapse = "|")))

#most common DEI words
rios_2w_count <- rios_data_token2it %>%
  filter(dei_related == "TRUE") %>%
  count(it_tokens_2w, sort = TRUE)


#graph of that 
rios_2w_count %>%
  top_n(30) %>%
  mutate(it_tokens_2w = reorder(it_tokens_2w, n)) %>%
  ggplot(aes(it_tokens_2w, n)) +
  geom_col() +
  coord_flip() +
  labs(y = "(DEI Related) 2 Word Count in Inclusive Teaching Text") + 
  xlab(NULL)


#total2dei for each year
total2deiyear <- rios_data_token2it %>%
  filter(dei_related == "TRUE") %>%
  group_by(Year) %>%
  count(it_tokens_2w, sort = TRUE) %>%
  summarise(totaldeiy = sum(n))

#totaldei for each article, only 240 obs vs 251!
total2deiarticle <- rios_data_token2it %>%
  filter(dei_related == "TRUE") %>%
  group_by(article_num) %>%
  count(it_tokens_2w, sort = TRUE) %>%
  summarise(totaldeia = sum(n)) %>%
  arrange(desc(totaldeia))

#total words in each year
totalwordsyear <- rios_data %>%
  group_by(Year) %>%
  summarise(totaly = sum(`Word Count of Inclusive Teaching?`, na.rm = TRUE))

#total words in each article
totalwordsarticle <- rios_data %>%
  select(15,9)
  

#proportions for each year, fluctuating like before
total2deiyear %>%
  full_join(totalwordsyear, by = c("Year")) %>%
  mutate(ratio = totaldeiy/totaly)%>%
  arrange(desc(ratio))

#proportions for each article, also fluctuating
total2deiarticle %>%
  full_join(totalwordsarticle, by = c("article_num")) %>%
  mutate(ratio = totaldeia/`Word Count of Inclusive Teaching?`) %>%
  arrange(desc(ratio))



```

## Network Plot of Word Relationship


```{r}
rios_data_token2 <- rios_data_token2it %>%
  separate(it_tokens_2w, c("word1", "word2"), sep = " ")

rios_data_token2 %>% 
  filter(dei_related == TRUE) %>%
  count(word1, word2, sort = TRUE) %>% 
  filter(n > 3) %>% 
  graph_from_data_frame()

```




